---
title: "MATH 242 - Final Project"
author: "Ziad and Phil"
date: "11/29/2017"
output:
  html_document: default
  word_document: default
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(Hmisc)
library(ggthemes)
library(knitr)
library(readr)
library(MASS)
```


```{r}
setwd("D:/College Study/MATH 242 - Applied Stats/Final Project")
EPL_Data<-read.csv("EPL_Data.csv", header = TRUE)

EPL_Data$position_cat <- factor(EPL_Data$position_cat,
                         levels = c(1,2,3,4),
                         labels = c("Attackers","Midfielders", "Defenders", "GoalKeepers"))

EPL_Data$region <- factor(EPL_Data$region,
                         levels = c(1,2,3,4),
                         labels = c("England","Europe", "America", "Rest of the World"))

EPL_Data$age_cat <- factor(EPL_Data$age_cat,
                         levels = c(1,2,3,4,5,6),
                         labels = c("18-21","22-24", "25-27", "28-31","32-34","35-above"))
```


# Explore Data


###Summary of players value and wage (by position groups)

```{r}
position_sum <- EPL_Data %>%
  group_by(position_cat) %>%
  summarise(Mean_Value = mean(eur_value), Min_Value = min(eur_value), Highest_Value = max(eur_value))


show(position_sum)
```

###Player Values Distribution by Position

```{r}
EPL_Data %>%
  ggplot(aes(x = position_cat, y = eur_value, colour = position_cat)) +
  geom_boxplot(alpha = 0.3, draw_quantiles = c(0.25, 0.5, 0.75)) + 
  geom_jitter(width = 0.2, height = 0, alpha = 0.2) +
  labs(x = "Position", y = "Market Value (euro)", colour = "Position", 
       title = "Player's Value Boxplots by Position") +
  theme_minimal()
```

###Summary of players value and wage (by regions)

```{r}
region_sum <- EPL_Data %>%
  group_by(region) %>%
  summarise(Mean_Value = mean(eur_value), Mean_Salary = mean(eur_wage), Highest_Value = max(eur_value))


show(region_sum)
```

###Player Values Distribution by Region

```{r}
EPL_Data %>%
  ggplot(aes(x = region, y = eur_value, colour = region)) +
  geom_boxplot() + 
  geom_jitter(width = 0.2, height = 0, alpha = 0.2) +
  labs(x = "Region", y = "Market Value (euro)", colour = "Region", 
       title = "Player's Value Distribution by Region") +
  theme_minimal()
```

###Player Values by Club and Age

```{r}
ageclub_sum <- EPL_Data %>%
  group_by(age_cat,big_club) %>%
  summarise(Mean_Value = mean(eur_value), Mean_Salary = mean(eur_wage), Highest_Value = max(eur_value))


show(ageclub_sum)
```

#Illustration of player' value by club and age

```{r}
EPL_Data %>% 
  group_by(age_cat, big_club) %>% 
  summarise(mean = mean(eur_value)) %>% 
  ggplot(aes(x = big_club, y = mean, fill = age_cat)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Big Clubs vs Small Clubs", y = "Average Value (euro)", fill = "Age Groups",
       title = "Comparison of Player's Value by Clubs and Age") +
  theme_minimal()
```

# Statistical Analysis


##1. Player's Value and Their Overall stat in FIFA

```{r}
EPL_Data %>% 
  ggplot(aes(overall,eur_value)) + geom_point() +
  labs(x = "Overall Stats", y = "Player's Value (euro)", title = "Relationship between Player's Value and Overall Stats") + theme(plot.title = element_text(hjust = 0.5)) + theme_tufte()
```

#Regression model with overall stats

```{r}
model1 = lm(sqrt(eur_value)~log(overall), EPL_Data)
summary(model1)
```

###Regression model with all the detailed stats

```{r}
model2 <- lm(log(eur_value)~pac + sho + pas + dri + def + pas, EPL_Data)
summary(model2)
```

###Create grouped data by positions

```{r}
#Group1
Group_Attack<-
  EPL_Data%>%
  filter(EPL_Data$position_cat == "Attackers")
#Group2
Group_Mid<-
  EPL_Data%>%
  filter(EPL_Data$position_cat == "Midfielders")
#Group3
Group_Defense<-
  EPL_Data%>%
  filter(EPL_Data$position_cat == "Defenders")
```

```{r}
regatt <- lm(log(eur_value)~ pac+sho+dri+pas+phy, Group_Attack)
summary(regatt)
```

```{r}
regcm <- lm(log(eur_value)~ sho+pas+dri+phy, Group_Mid)
summary(regcm)
```

```{r}
regdef <- lm(log(eur_value)~ def+phy+pas, Group_Defense)
summary(regdef)
```

##Player's Value vs Age

```{r}
EPL_Data %>% 
  ggplot(aes(age,eur_value)) + geom_jitter() +
  labs(x = "Age", y = "Player's Value (euro)", title = "Relationship between Player's Value and Age") + theme(plot.title = element_text(hjust = 0.5)) + theme_minimal()
```

```{r}
EPL_Data$agesq = EPL_Data$age*EPL_Data$age
reg_age <- lm(sqrt(eur_value)~age+agesq, EPL_Data)
summary(reg_age)
```

##Player's Value vs Popularity

```{r}
EPL_Data %>% 
  ggplot(aes(sqrt(page_views),sqrt(eur_value))) + geom_jitter() +
  labs(x = "Page Views", y = "Player's Value (euro)", title = "Relationship between Player's Value and Popularity") + theme(plot.title = element_text(hjust = 0.5)) + theme_minimal()
```


###Player's value vs Fpl_point

```{r}
EPL_Data %>% 
  ggplot(aes(fpl_value,eur_value)) + geom_jitter() +
  labs(x = "Fantasy Premier League value", y = "Player's Value (euro)", title = "Relationship between Player's Value and Their Fantasy League Value") + theme(plot.title = element_text(hjust = 0.5)) + theme_minimal()
```

```{r}
EPL_Data %>% 
  ggplot(aes(eur_wage,eur_value)) + geom_jitter() +
  labs(x = "Player's Wage", y = "Player's Value (euro)", title = "Relationship between Player's Value and Their Wage") + theme(plot.title = element_text(hjust = 0.5)) + theme_minimal()
```

```{r}
EPL_Data %>% 
  ggplot(aes(international_reputation,eur_value)) + geom_jitter() +
  labs(x = "International Reputation", y = "Player's Value (euro)", title = "Relationship between Player's Value and Their international Reputation") + theme(plot.title = element_text(hjust = 0.5)) + theme_minimal()
```


```{r}
step = stepAIC(lm(sqrt(eur_value)~1,data=EPL_Data),scope=~ sqrt(eur_wage) + sqrt(page_views) + log(overall) + sho + dri + def + fpl_value + age + agesq + fpl_points + as.factor(region) + new_foreign + new_signing + as.factor(preferred_foot) + big_club + as.factor(position_cat) + page_views*new_signing + international_reputation*new_signing + height_cm*weight_kg, direction="forward")
step$anova # display summarized results
```


```{r}
modeltest <- lm(sqrt(eur_value) ~ sqrt(eur_wage) + log(overall) + age + agesq + fpl_value + international_reputation + as.factor(position_cat) + big_club + sho, EPL_Data)

summary(modeltest)
```


```{r}
bestmodel <- lm(sqrt(eur_value) ~ eur_wage + log(overall) + age + agesq + fpl_value + 
    as.factor(position_cat) + big_club, EPL_Data)

summary(bestmodel)
```


###Assumption1:

```{r}
residual1 <- EPL_Data %>% mutate(resid1 = resid(bestmodel))

residual1 %>%
  ggplot(aes(eur_wage, resid1)) +
  geom_point(colour = "blue") + geom_hline(yintercept = 0)
```


###Assumption 2:
```{r}
resbm <- bestmodel$res
qqnorm(resbm)
qqline(resbm)
```



```{r}
EPL_Data$Predicted <- fitted(bestmodel)^2
EPL_Data$Residual <- EPL_Data$eur_value - EPL_Data$Predicted


table_value <- EPL_Data %>% filter(big_club ==1)

table_value <- subset(table_value, select=c(name,club,eur_value,Predicted,Residual, fpl_points))
```

```{r}
#quantile(table_value$Predicted)

summary(table_value$Residual)
```

```{r}
overvalued_player <- table_value %>% filter(Residual > 2.273)

overvalued_player <- subset(overvalued_player, select=c(name,club,eur_value,Predicted,fpl_points, Residual))

overvalued_player
```


```{r}
value_sum <- table_value %>%
  group_by(club) %>%
  summarise(Overvalued = sum(Residual > 2.273))


show(value_sum)
```

```{r}
overvalued_percentage <- table_value %>%
  group_by(club) %>%
  summarise(Percentage = sum(fpl_points[Residual> 2.273])/sum(fpl_points))

overvalued_percentage
```













